### 2024-04-04 paper/paper.qmd
update change

---
title: "Demographic Groups with Stronger Biden Support in the Election Race'"
author: 
  - Yang Zhou
thanks: "Code and data are available at: <https://github.com/yangzhoucoco/Political-support-in-the-United-State.git >"
date: today
date-format: long
abstract: "This paper explores the relationship between voting preferences and the predictors of education, income, and personal gun ownership in the United States through analyzing 2022 US election database. By examining data from Cooperative Election Study, we identify clear patterns indicating that higher income are more likely to vote for Biden. Besides, the higher personal gun ownership will also significantly increase the likelihood of voting for Biden. By contrast, the people with higher education level show less probability to vote for Biden. Our findings underscore the critical impact of socio-economic status and personal security concerns on electoral outcomes, shedding light on the underlying dynamics that shape voter behavior in contemporary American politics. Ultimately, this research contributes to our understanding of the complex factors that drive electoral decisions, offering valuable insights for policymakers, political strategists, and citizens aiming to foster more informed and equitable democratic processes."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(modelsummary)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
```

# Introduction
2020 United States Presidential Election stood as an important political event, reflecting societal divisions and highlighting the influence of various demographic factors on voting behavior. This election, which culminated in the victory of Joe Biden over incumbent Donald Trump, has spurred a renewed interest in understanding the dynamics of voter preferences and the underlying factors that drive electoral decisions. Our study aims to dissect the relationship between voters' educational background, income levels, personal gun ownership, and their voting preferences, employing logistic regression analysis to reveal the potential associations between different factors.

The Cooperative Election Study (CES)conduct a throughout survey encompassing a wide range of voter demographics and attitudes.CES collaborate with 60 researches teams to assembles sample comprising 60,000 cases. The cases are based on the survey of adult pariticipants in 2022 fall. YouGov was commissioned to help the teams to conduct national sample surbey. They conducted the survey before and after elections. The first round survey were completed from September 29 to November 8, 2022 and the second period is gathered from November 10 to December 15, 2022. By offering an throughout views to different constituencies, voter demographics and voter behaviors, CES reveals the complexity of electoral behavior and voter demographics. By understanding voter behaviors,, we gain valuable insights that pave the way for refining and advancing democratic practices.

In addition, by leveraging data from the 2022 CES, this paper delves into how variables such as education, income, and gun ownership serve as predictors for electoral choices, specifically the likelihood of voting for Biden or Trump. This analysis not only reveal the individual impact of these factors but also contributes to a broader understanding of the socio-economic and cultural impact of American electoral behavior.

Based on our model, I found that individuals with higher incomes demonstrate a greater propensity to vote for Biden. Additionally, elevated levels of personal gun ownership significantly correlate with an increased likelihood of supporting Biden. Conversely, individuals possessing higher levels of education exhibit a reduced probability of voting for Biden.

To be more specific, this paper is structured as follow: in {#sec-data}, I will introduce the data variable and how the data is used for the overall analysis.Visualization would also be included. In {#sec-model}, I will set up the model to predict the relationship between the predictors and electoral outcomes. Model justification will provide a rationale for selecting logitics model. {#sec-results} is focused on explain the result of the model. Lastly, {#sec-Discussion} provides a discussion on what we found in the model. I will also talk about the weakness of this paper along with the further study on this topic.

I will uses the programming language `R`[@citeR] to analyze the data. `dplyr`[@dplyr],  `tidyverse`[@tidyverse], `ggplot2`[@ggplot], `knitr` [@knitr],  `here` [@here] , and `kableExtra` [@kableExtra],`modelsummary`[@model_summary], and `rstanarm`[@rstanarm] will help me in the model and visualization part.

# Data {#sec-data}

## Source and Methodology
The 2022 Cooperative Election Study (CES) [@main_data] was conducted by YouGov. In order to ensure the representativesness of the data, the survey utilized sampling methodology, matching process, weighting procedure.

## Sampling and Sampling Matching
The survey interviewed 60,000 adults during two periods: pre-election from September 29 to November 8, 2022, and post-election from November 10 to December 15, 2022. YouGov employed matched random sample methodology for the sampling process. Sample matching is used to create representative samples from non-random pools of respondents and it is suitable for online survey. First,it will draw a random target population and then, it is going to select the matching respondents from the pool which matched with the target sample's characteristics. In order to replicate the target sample's attributes, YouGov uses proximity matching to calculates the likelihood or closeness between characteristics of the target samples; they will adjust the weight of variables when they need.

## Weightening
To correct the imbalance between the matched samples and the overarching target demographic, a two-phase weighting procedure is employed, ensuring the CES samples accurately mirror the population's diversity. The initial phase of this adjustment process involves entropy balancing, alongside iterative proportional fitting (often termed 'raking'), to align the sample with the population's demographic and political characteristics. This includes a comprehensive evaluation of variables and their interrelations. The approach guarantees that the common content of the survey represents each state accurately, incorporating adjustments for statewide electoral contests.

## Vote Validation
Vote validation have been considered in the CES in order to verify the accuracy of the voting behaviors. The sample were matched to the TargetSmart database that include the registered voter responds. Only the records which have a high level of confidence could be matched. This process is helpful for identifying voters and the way they vote, such as absentee, early, and mail.Therefore CES can ensure the data is focused on voting behavior and accurately represent genuine voter engagement.

## Variables
```{r}
#| label: fig-variables
#| fig-cap: Brief summary of variables types
#| echo: false
#| warning: false
#| message: false

ces2022 <-  read_csv(file = here::here("data/raw_datas/raw_datas.csv"))
    col_types =
      cols(
        "votereg" = col_integer(),
        "presvote20post" = col_integer(),
        "educ" = col_integer(),
        "faminc_new" = col_integer(),
        "gunown" = col_integer()
      )

ces2022 |> slice_head(n = 10) |>
  kable(booktabs = TRUE, caption = "Brief summary of variables types") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```
After installing `dataverse`[@dataverse], I use `get_dataframe_by_name` to analyze CES. I select four variables that most relevant to our topic, which are `votereg`, `presvote20post`, `educ`, `faminc_new`, `gunown`.  `votereg` represents the voter registration status and only the voter who registered to vote would be considered. Besides, `presvote20post` represents the president that respondents vote in 2020 president election. `educ` means the education level of the respondents. In addition, `faminc_new` accounts for the family's annual income in 2021 year. `gunown` records the personal gun ownership by asking the respondent to answer whether they or anyone in their household own a gun. However, for the raw data, there are six different answers, including choosing ,Joe Biden, Donald Trump, Jo Jorgensen, Howie Hawkins, other, and did not vote for President. I only consider the respond of Joe Biden and Donald Trump. Therefore, when I clean the data, I select the group of Joe Biden and Donald Trump. In addition, there are 6 different education level, ranging from no high school, high school graduate, some college, 2-year, 4-year, and post-grad. I care about all of the education level, so I take all of those types into consideration. Moreover, there are 16 ranges of family income and the differences between each range is 10,000. To simiplied the data, I classifed them into 4 different types. For the income group which Less than $10,000,  $10,000 - $19,999, $20,000 - $29,999, and $30,000 - $39,999 are defined as "Low" income group. For those who range from $40,000 to $79,999 can seem as  "Middle" income group.  $80,000 to $199,999 can be considered as "High" income group. For the family earn $200,000 to $500,000 or more is very high income group.

```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and gun ownership
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

ggplot(ces2022, aes(x = income_group, fill = voted_for)) +  
  geom_bar(position = "dodge") +  
  geom_text(
    aes(label = ..count..),
    stat = 'count',
    position = position_dodge(width = 0.9),
    vjust = -0.25,
    color = "black",
    size = 2.5
  ) +
  labs(
    x = "Income Level",  
    y = "Number of Respondents",  
    fill = "Voted for"
  ) +
  coord_flip() +  
  scale_fill_brewer(palette = "Set1") +  
  theme_minimal(base_size = 14) +  
  theme(
    legend.position = "bottom", 
    strip.text = element_text(size = 12)
  )

```

```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and gun ownership
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

ggplot(ces2022, aes(x = educ, fill = voted_for)) +  
  geom_bar(position = "dodge") +  
  geom_text(
    aes(label = ..count..),
    stat = 'count',
    position = position_dodge(width = 0.9),
    vjust = -0.25,
    color = "black",
    size = 2.5
  ) +
  labs(
    x = "Education Level",  
    y = "Number of Respondents",  
    fill = "Voted for"
  ) +
  coord_flip() +  
  scale_fill_brewer(palette = "Set1") +  
  theme_minimal(base_size = 14) +  
  theme(
    legend.position = "bottom", 
    strip.text = element_text(size = 12)
  )

```
```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and gun ownership
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

# Generate the plot focusing on Personal Gun Ownership
ggplot(ces2022, aes(x = gunown, fill = voted_for)) +  
  geom_bar(position = "dodge") +  
  geom_text(
    aes(label = ..count..),
    stat = 'count',
    position = position_dodge(width = 0.9),
    vjust = -0.25,
    color = "black",
    size = 2.5
  ) +
  labs(
    x = "Personal Gun Ownership",  
    y = "Number of Respondents",  
    fill = "Voted for"
  ) +
  coord_flip() +  
  scale_fill_brewer(palette = "Set1") +  
  theme_minimal(base_size = 14) +  
  theme(
    legend.position = "bottom", 
    strip.text = element_text(size = 12)
  )

```

```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and income level
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

ggplot(ces2022, aes(x = educ, fill = voted_for)) +
  geom_bar(position = "dodge") +  
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.25, position = position_dodge(0.9), color = "black", size = 2.5) +
  facet_wrap(~income_group) +  
  labs(x = "Education Level", y = "Number of Respondents", fill = "Vote") +
  coord_flip() +  
  theme_minimal() +  
  scale_fill_brewer(palette = "Set1") +  
  theme(legend.position = "bottom", strip.text = element_text(size = 12))  
```

```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and income level
#| echo: false
#| warning: false
#| message: false

ggplot(ces2022, aes(x = income_group, fill = voted_for)) +
  geom_bar(position = "fill") +
  labs(x = "Income Group", y = "Proportion", fill = "Voted For") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

## Measurements

# Model {#sec-model}
## Model set-up

Define $y_i$ is the political preference of the respondent, where $y_i$ equal to 1 indicates a preference for Biden and $y_i$ equal to 0 represent a preference for Trump. The predictors include $\mbox{education}_i$, which indicates the education level of the respondent and $\mbox{income}_i$ is the income level of the respondent. $\mbox{gunownership}_i$ is the state of whether individual or the people he/she knows own a gun.

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta \times \mbox{education}_i + \gamma \times \mbox{income}_i + \delta \times \mbox{gun}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\delta &\sim \mbox{Normal}(0, 2.5)
\end{align}

I set up a logistic regression model within a Bayesian framework to estimate model parameters, which particularly focusing on binary outcome variables such as political preference. To be more specific, I utilize `stan_glm()` function from `rstanrm` [@rstanrm] package, which can handles categorical variables efficiently by converting them into binary indicators. In order to optimize computational efficiency, I use a random sample of 500 observations to fit the model.

## Model justification
Logistic regression is designed for binary outcome variables,especially when the dependent variables indicates two categories, in this case, the dependent variables are voting for Biden(1) or Trump(0). This model can provide a straightforward method for estimating the probability for voting for a specific candidate based on its predictors, including education level, income level, and gun ownership. In addition, Logistic regression is an ideal choice for explaining how unit changes in predictors variables affect the dependent variables.

# Results {#sec-results}

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

us_political_preferences <-
  readRDS(file = here::here("model/first_model.rds"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-model-summary-table
#| tbl-cap: "Explanatory models of political preferences based on gender and race (n = 500)"

modelsummary(
  list(
    "Support Biden" = us_political_preferences
  ),
  statistic = "mad"
)

```

# Discussion {#sec-Discussion}

## Findings

## High income

## Education

## Gun ownership

## Application of Logistic Regression

## Linear Regression Model

## Model Building

## Weaknesses and Next Steps

\newpage

# References {#sec-Reference}