### 2024-04-05 paper/paper.qmd
update change

---
title: "Demographic Groups with Stronger Biden Support in the Election Race'"
author: 
  - Yang Zhou
thanks: "Code and data are available at: <https://github.com/yangzhoucoco/Political-support-in-the-United-State.git >"
date: today
date-format: long
abstract: "This paper explores the relationship between voting preferences and the predictors of education, income, and personal gun ownership in the United States through analyzing 2022 US election database. By examining data from Cooperative Election Study, we identify clear patterns indicating that higher income are more likely to vote for Biden. Besides, the higher personal gun ownership will also significantly increase the likelihood of voting for Biden. By contrast, the people with higher education level show less probability to vote for Biden. Our findings underscore the critical impact of socio-economic status and personal security concerns on electoral outcomes, shedding light on the underlying dynamics that shape voter behavior in contemporary American politics. Ultimately, this research contributes to our understanding of the complex factors that drive electoral decisions, offering valuable insights for policymakers, political strategists, and citizens aiming to foster more informed and equitable democratic processes."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(modelsummary)
library(knitr)
library(kableExtra)
library(readr)
library(ggplot2)
library(rstanarm)
```

# Introduction
2020 United States Presidential Election stood as an important political event, reflecting societal divisions and highlighting the influence of various demographic factors on voting behavior. This election, which culminated in the victory of Joe Biden over incumbent Donald Trump, has spurred a renewed interest in understanding the dynamics of voter preferences and the underlying factors that drive electoral decisions. Our study aims to dissect the relationship between voters' educational background, income levels, personal gun ownership, and their voting preferences, employing logistic regression analysis to reveal the potential associations between different factors.

The Cooperative Election Study (CES)conduct a throughout survey encompassing a wide range of voter demographics and attitudes.CES collaborate with 60 researches teams to assembles sample comprising 60,000 cases. The cases are based on the survey of adult pariticipants in 2022 fall. YouGov was commissioned to help the teams to conduct national sample surbey. They conducted the survey before and after elections. The first round survey were completed from September 29 to November 8, 2022 and the second period is gathered from November 10 to December 15, 2022. By offering an throughout views to different constituencies, voter demographics and voter behaviors, CES reveals the complexity of electoral behavior and voter demographics. By understanding voter behaviors,, we gain valuable insights that pave the way for refining and advancing democratic practices.

In addition, by leveraging data from the 2022 CES, this paper delves into how variables such as education, income, and gun ownership serve as predictors for electoral choices, specifically the likelihood of voting for Biden or Trump. This analysis not only reveal the individual impact of these factors but also contributes to a broader understanding of the socio-economic and cultural impact of American electoral behavior.

Based on our model, I found that individuals with higher incomes demonstrate a greater propensity to vote for Biden. Additionally, elevated levels of personal gun ownership significantly correlate with an increased likelihood of supporting Biden. Conversely, individuals possessing higher levels of education exhibit a reduced probability of voting for Biden.

To be more specific, this paper is structured as follow: in {#sec-data}, I will introduce the data variable and how the data is used for the overall analysis.Visualization would also be included. In {#sec-model}, I will set up the model to predict the relationship between the predictors and electoral outcomes. Model justification will provide a rationale for selecting logitics model. {#sec-results} is focused on explain the result of the model. Lastly, {#sec-Discussion} provides a discussion on what we found in the model. I will also talk about the weakness of this paper along with the further study on this topic.

I will uses the programming language `R`[@citeR] to analyze the data. `dplyr`[@dplyr],  `tidyverse`[@tidyverse], `ggplot2`[@ggplot], `knitr` [@knitr],  `here` [@here] , and `kableExtra` [@kableExtra],`modelsummary`[@model_summary], and `rstanarm`[@rstanarm] will help me in the model and visualization part.

# Data {#sec-data}

## Source
The 2022 Cooperative Election Study (CES) [@main_data] was conducted by YouGov. In order to ensure the representativesness of the data, the survey utilized sampling methodology, matching process, weighting procedure.

## Sampling
The survey interviewed 60,000 adults during two periods: pre-election from September 29 to November 8, 2022, and post-election from November 10 to December 15, 2022. YouGov employed matched random sample methodology for the sampling process. Sample matching is used to create representative samples from non-random pools of respondents and it is suitable for online survey. First,it will draw a random target population and then, it is going to select the matching respondents from the pool which matched with the target sample's characteristics. In order to replicate the target sample's attributes, YouGov uses proximity matching to calculates the likelihood or closeness between characteristics of the target samples; they will adjust the weight of variables when they need.

## Weightening
To correct the imbalance between the matched samples and the overarching target demographic, a two-phase weighting procedure is employed, ensuring the CES samples accurately mirror the population's diversity. The initial phase of this adjustment process involves entropy balancing, alongside iterative proportional fitting (often termed 'raking'), to align the sample with the population's demographic and political characteristics. This includes a comprehensive evaluation of variables and their interrelations. The approach guarantees that the common content of the survey represents each state accurately, incorporating adjustments for statewide electoral contests.

## Vote Validation
Vote validation have been considered in the CES in order to verify the accuracy of the voting behaviors. The sample were matched to the TargetSmart database that include the registered voter responds. Only the records which have a high level of confidence could be matched. This process is helpful for identifying voters and the way they vote, such as absentee, early, and mail.Therefore CES can ensure the data is focused on voting behavior and accurately represent genuine voter engagement.

## Variables
```{r}
#| label: fig-variables
#| fig-cap: Brief summary of variables types
#| echo: false
#| warning: false
#| message: false

ces2022 <-  read_csv(file = here::here("data/raw_datas/raw_datas.csv"))
    col_types =
      cols(
        "votereg" = col_integer(),
        "presvote20post" = col_integer(),
        "educ" = col_integer(),
        "faminc_new" = col_integer(),
        "gunown" = col_integer()
      )

ces2022 |> slice_head(n = 10) |>
  kable(booktabs = TRUE, caption = "Brief summary of variables types") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)
```
After installing `dataverse`[@dataverse], I use `get_dataframe_by_name` to analyze CES. I select four variables that most relevant to our topic, which are `votereg`, `presvote20post`, `educ`, `faminc_new`, `gunown`.  `votereg` represents the voter registration status and only the voter who registered to vote would be considered. Besides, `presvote20post` represents the president that respondents vote in 2020 president election. `educ` means the education level of the respondents. In addition, `faminc_new` accounts for the family's annual income in 2021 year. `gunown` records the personal gun ownership by asking the respondent to answer whether they or anyone in their household own a gun. However, for the raw data, there are six different answers, including choosing ,Joe Biden, Donald Trump, Jo Jorgensen, Howie Hawkins, other, and did not vote for President. I only consider the respond of Joe Biden and Donald Trump. Therefore, when I clean the data, I select the group of Joe Biden and Donald Trump. In addition, there are 6 different education level, ranging from no high school, high school graduate, some college, 2-year, 4-year, and post-grad. I care about all of the education level, so I take all of those types into consideration. Moreover, there are 16 ranges of family income and the differences between each range is 10,000. To simiplied the data, I classifed them into 4 different types. For the income group which Less than $10,000,  $10,000 - $19,999, $20,000 - $29,999, and $30,000 - $39,999 are defined as "Low" income group. For those who range from $40,000 to $79,999 can seem as  "Middle" income group.  $80,000 to $199,999 can be considered as "High" income group. For the family earn $200,000 to $500,000 or more is very high income group.

```{r}
#| label: fig-income-vote
#| fig-cap: The distribution of presidential preferences, by income level
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

ggplot(ces2022, aes(x = income_group, fill = voted_for)) +  
  geom_bar(position = "dodge") +  
  geom_text(
    aes(label = ..count..),
    stat = 'count',
    position = position_dodge(width = 0.9),
    vjust = -0.25,
    color = "black",
    size = 2.5
  ) +
  labs(
    x = "Income Level",  
    y = "Number of Respondents",  
    fill = "Voted for"
  ) +
  coord_flip() +  
  scale_fill_brewer(palette = "Set1") +  
  theme_minimal(base_size = 14) +  
  theme(
    legend.position = "bottom", 
    strip.text = element_text(size = 12)
  )

```

```{r}
#| label: fig-gun-ownership-vote
#| fig-cap: The distribution of presidential preferences, by gun ownership
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

# Generate the plot focusing on Personal Gun Ownership
ggplot(ces2022, aes(x = gunown, fill = voted_for)) +  
  geom_bar(position = "dodge") +  
  geom_text(
    aes(label = ..count..),
    stat = 'count',
    position = position_dodge(width = 0.9),
    vjust = -0.25,
    color = "black",
    size = 2.5
  ) +
  labs(
    x = "Personal Gun Ownership",  
    y = "Number of Respondents",  
    fill = "Voted for"
  ) +
  coord_flip() +  
  scale_fill_brewer(palette = "Set1") +  
  theme_minimal(base_size = 14) +  
  theme(
    legend.position = "bottom", 
    strip.text = element_text(size = 12)
  )

```

```{r}
#| label: fig-education-income-vote
#| fig-cap: The distribution of presidential preferences, by education, and income level
#| echo: false
#| warning: false
#| message: false

ces2022 <- read_csv(here("data/analysis_data/analysis_data.csv"))

ggplot(ces2022, aes(x = educ, fill = voted_for)) +
  geom_bar(position = "dodge") +  
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.25, position = position_dodge(0.9), color = "black", size = 2.5) +
  facet_wrap(~income_group) +  
  labs(x = "Education Level", y = "Number of Respondents", fill = "Vote") +
  coord_flip() +  
  theme_minimal() +  
  scale_fill_brewer(palette = "Set1") +  
  theme(legend.position = "bottom", strip.text = element_text(size = 12))  
```

```{r}
#| label: fig-vote
#| fig-cap: The distribution of presidential preferences, by education, and income level
#| echo: false
#| warning: false
#| message: false

ggplot(ces2022, aes(x = income_group, fill = voted_for)) +
  geom_bar(position = "fill") +
  labs(x = "Income Group", y = "Proportion", fill = "Voted For") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

## Measurements

# Model {#sec-model}
## Model set-up

Define $y_i$ is the political preference of the respondent, where $y_i$ equal to 1 indicates a preference for Biden and $y_i$ equal to 0 represent a preference for Trump. The predictors include $\mbox{education}_i$, which indicates the education level of the respondent and $\mbox{income}_i$ is the income level of the respondent. $\mbox{gunownership}_i$ is the state of whether individual or the people he/she knows own a gun.

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta \times \mbox{education}_i + \gamma \times \mbox{income}_i + \delta \times \mbox{gun}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\delta &\sim \mbox{Normal}(0, 2.5)
\end{align}

I set up a logistic regression model within a Bayesian framework to estimate model parameters, which particularly focusing on binary outcome variables such as political preference. To be more specific, I utilize `stan_glm()` function from `rstanrm` [@rstanrm] package, which can handles categorical variables efficiently by converting them into binary indicators. In order to optimize computational efficiency, I use a random sample of 500 observations to fit the model.

## Model justification
Logistic regression is designed for binary outcome variables,especially when the dependent variables indicates two categories, in this case, the dependent variables are voting for Biden(1) or Trump(0). This model can provide a straightforward method for estimating the probability for voting for a specific candidate based on its predictors, including education level, income level, and gun ownership. In addition, Logistic regression is an ideal choice for explaining how unit changes in predictors variables affect the dependent variables.

# Results {#sec-results}
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

voting_preference <-
  readRDS(file = here::here("model/first_model.rds"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: tbl-model-summary-table
#| tbl-cap: "Explanatory models of political preferences based on education, income and gun ownership (n = 500)"

modelsummary(
  list(
    "Support Biden" = voting_preference
  ),
  statistic = "mad"
)

```
After conducting a logistic regression analysis on 500 samples, I summary the result in @tbl-model-summary-table. According to @Mustafa_2024, the coefficients of the model are referred to posterior means which provides a central tendency measure of parameter's posterior distribution. 
The coefficient represents the log odds of supporting Biden for the reference group while all other variables are held at their baseline values.
I care about three predictors which are education level, income level, and gun ownership and focus on the relationship of political preference of different groups and their reference group. To be more specific, in terms of education level, the model interpret the relationship between the voters who are 4-year education, high school graduate, no high school, post graduate, some college and the reference group with a 2-year education. The reference group for income level is high income group range from $80,000 to $199,999 and its compared to other income group, including low income, middles income, and very high income. For the individual gun ownership, "no one in the household owns a gun" is the reference group.
For example, the coefficient for voters with a 4-year education level is -0.251. This indicates that holding all other variables constant, the voters having a 4-year education compared to the reference group(2-year education) is associated with a slight decreases.

## Education
According to @tbl-model-summary-table, the respondents who have high-school degree(1.337), no high school degree(1.131), and voter have some college degree (0.482) are more likely to vote for Biden, whereas, the people with 4-year education(-0.251), or post-graduate degree (-0.108) are less likely to vote for Biden.

## Income
The respondent with low income, ranging from less than $10,000 to $39,999, and very high income group, ranging from $200,000 to $500,000 or more, are less likely to vote for Biden. Their coefficients are −0.207 and −0.954 respectively. The middle income group is more likely to vote for Biden, ranging from $40,000 to $79,999.

## Gun ownership
People own a gun(1.058) or know someone in the household owns a gun(0.280) have a larger likelihood to vote for Biden.

# Discussion {#sec-Discussion}
## Findings
After I used a Bayesian logistic regression model to find the relationship between Biden support and the predictors, including education level, income level, and individual gun ownership, I conclude three main findings shown below:
  1. People have higher education level is less likely to vote for Biden than those of lower education level
  2. Middle income group tend to vote for Biden
  3. The voters who own gun or know someone who own gun exhibit a higher propensity to vote for Biden

## People have higher education level is less likely to vote for Biden than those of lower education level
The summary data presented in @tbl-model-summary-table reveals a notable electoral dynamic: individuals possessing a four-year undergraduate degree or higher education credentials exhibit a lower propensity to support President Biden, in contrast to those whose highest educational attainment is a high school diploma or below, who show a marked tendency towards endorsing him. This pattern suggests that educational attainment may play a significant role in shaping political preferences, though it's important to acknowledge that a multitude of factors can influence an individual's voting decisions.

A critical consideration in unpacking these trends is the tax policy proposed by President Biden's administration, as outlined by @The_White_House_2024. The policy seeks to increase the marginal income tax rate to 39.6% for individuals whose annual earnings exceed $400,000. Notably, this income bracket is more likely to include individuals with advanced degrees. The people often command higher salaries because of their education level. Thus, the observed correlation between higher educational levels and diminished support for Biden might lead to considerations within this high earner regarding the potential impact of the proposed tax adjustments on their financial well-being.

Such concerns are not merely about the taxation but also about the perceived fairness and implications of these tax changes. High-earning individuals may worry about bearing a disproportionate tax burden. Additionally, this group might conduct the survey on the long-term economic impacts of higher taxes on investment, job creation, and overall economic growth.

Furthermore, this discussion highlights the complexity between education, income, and political ideology. While higher education often correlates with higher income, it also fosters exposure to diverse ideas and critical thinking skills that can influence political beliefs and policy preferences. Therefore, the relationship between education, income, and political preference might be complex.

## Middle income group tend to vote for Biden
The analysis presented in @fig-income-vote compellingly demonstrates that Biden’s support base comprises significantly of middle and high-income groups, with the middle-income segment being particularly dominate. This trend is substantiated by data in @tbl-model-summary-table, which indicates positive correlation between support for Biden and belonging to the middle-income group. While there is no a specific analyses explicating the motivations behind this pattern, it is plausible to attribute this trend to Biden’s policy that directly benefit middle-income families, such as the enhancement of tax credits. According to @Parys, Biden has supported expanding tax credits for middle-income families, such as Children Tax Credit, which a measure designed to alleviate the financial burden of child-rearing for families. Middle income group could benefit from this policy. Under Biden's administration, according to @US_Department_of_the_Treasury_2022, there was an increased credit amount in 2021; the credit amount raised from $2,000 each child to $3,600 for each child who aged from 6 to 17. Unlike previous version, Biden support a fully refundable credit policy which aim to provide a direct financial assistance to families. 
This policy is really about helping middle-income families by giving them tax breaks and other benefits. The Biden administration is focusing on these families to help them with their money problems, which is why many of them support Biden.

## The voters who own gun or know someone who own gun exhibit a higher propensity to vote for Biden 
The data presented in @tbl-model-summary-table reveals an interesting trend: individuals who own guns or are acquainted with gun owners exhibit a higher likelihood of supporting President Biden, despite he advocates more stringent gun control measures. Biden seeks to implement universal background checks, which called Bipartisan Background Checks Act. @Bipartisan_Background_Checks_Act aims to expand background check to cover all sales and transfers of weapons, including private transaction, online sales, and gun show sales.

This apparent paradox, where gun owners show support for a candidate known for promoting restrictive gun policies. It is plausible that gun owners, while cherishing their  rights, are equally concerned about the pervasive issue of gun violence concerns in the nation. Their support for Biden may because they recognize the measures such as universal background checks represent a balanced approach to regulate gun violence without infringing on the legitimate rights of the citizens to own firearms for personal security and recreational purposes.

## Weaknesses and Next Steps

\newpage

# References {#sec-Reference}